\documentclass[journal]{IEEEtran}
%\documentclass[conference]{IEEEtran}
%\documentclass[a4paper, 11pt]{article}

\usepackage[english]{babel}
\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage{url}
\usepackage{xspace}
\usepackage{eurosym}
\usepackage{lmodern} % for textsc

\usepackage{authblk} % for authors affiliation

\usepackage{hyperref}
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = blue, %Colour of citations
  breaklinks = true %use \href{url}{\nolinkurl{display URL}} for breakable long URLs
}

% \topmargin = 0pt
% \headheight = 5pt
% \headsep = 10pt
% \textwidth = 410pt
% \textheight = 650pt
% \oddsidemargin = 20pt

\title{A café-philo for computer science:\\motives, experience and next steps}
\date{}

\author[1,3]{Marc Bruyère}
\author[2,3]{Florian Richoux}
\author[4]{Daphné Tuncer}
\author[5]{Marc van der Wal}
\affil[   ]{\it   marc@iij.ad.jp    \quad   florian@richoux.fr   \quad
  daphne.tuncer@enpc.fr \quad marc.vanderwal@afnic.fr}
\affil[1]{IIJ Research Lab, Japan}
\affil[2]{AIST, Japan}
\affil[3]{JFLI, CNRS, Japan}
\affil[4]{École des Ponts ParisTech, France}
\affil[5]{Afnic, France}

\begin{document}
\maketitle

\begin{abstract}
  In  this short  paper,  we report  on our  experience  setting up  a
  monthly  reading group  around  the  socio-technical challenges  and
  ethics of computer science and  digitalization. Our reading group is
  the result  of an  encounter between  four computer  scientists with
  different backgrounds and career paths,  who work in different areas
  of computer systems, and who  live in different places. Beyond these
  factual differences,  we share an  eagerness to talk with  our peers
  not only  about our technical contributions  to digital developments
  but also about their implications on the future of our societies.
\end{abstract}

\section{The need to discuss}

The  incidence of  digital developments  in day-to-day  life has  been
massive over  the past  30 years.  From the  way a  large part  of our
interactions  takes  place  today  to how vital  infrastructures
operate or the  world’s geopolitics is shaped, the realization
of digital tools and technologies is far from being a technical matter
only. As practitioners of computer science, both through our education
in college and through our  professional occupation, we, as co-authors
of this paper, acknowledge that  we have directly been contributing to
the construction of this digital era.

Before being computer scientists, we are people who live, work, and evolve in
a set era and a set system, and indifferently from non computer scientists,
benefit from the outcomes of these digital developments. This dual status of
operators and beneficiaries of technological developments makes us ponder upon
the emergent effects of the use and applications of digital constructs and
technologies in the wild. It is however not always straightforward to
communicate these interrogations within our professional communities that
often dismiss such considerations as irrelevant to their purpose.

Also, we do realize that talking with our peers about these issues
is fundamental for broadening a perspective that can otherwise easily be
defined  by  the  view  of   a tiny fraction   of  the  world’s
population. We  therefore decided  to set  a reading  group up  in the
manner of  a \textit{café-philo}  for computer  science that  lets participants
share and  express their  opinions on anything  related to  the digital
world. Our discussions span many topics, such as the possibilities and
opportunities of AI for people, but also the misuse of AI
for purposes that are detrimental to the common good, the digital divide, the
society of control, and the undeniable environmental implications and
sustainability issues of widespread digitalization.

\section{Reading to debate}

For the four of us, the reading group is a side project. It takes place in
sessions organized on a monthly basis. The setting is informal. The objective
is to encourage the free sharing of ideas and opinions. Discussions are
only constrained by the one-hour time limit reserved for this activity.

Each  session  is led  by  one  of  the  participants, on  a  rotating
basis. The  lead for the  session presents a  piece of material  to be
discussed. The choice  of the material is entirely up  to the lead. It
can be  a scientific publication,  a book  targeted to a  more general
audience,  an audio clip, a video,  a documentary,  an  article from  a
newspaper, \textit{etc.}; anything that can  trigger a discussion regarding the
role and impact  of digitalization. Sessions are  usually not formally
concluded. The debate is left open  and the topics discussed are aimed
to support each of us with new ideas and perspectives to be applied in
our practice, as well as to consolidate a general reflection.

Our reading  group has been  meeting for  two years during  which many
different topics  were debated. One  of the sessions focused for instance
on the workshop organized   by  the   Internet Architecture Board in 2022
on the environmental  impact of Internet   applications  and   systems \cite{IAB22}.
Other sessions discussed different AI developements, \textit{e.g.,} with the recent
paper by Michael  Cook on the social responsibility of AI applied to video games \cite{Cook21},
the  book \textit{Weapons  of  Math Destruction} by  Cathy O’Neil  \cite{WMD16} or  a selection  of papers
presented at the ACM FAccT 2023 conference \cite{bell2023_fairness,karan2023_differential_pricing,lucaj2023_ai_regulation,rao2023_facebook_discrimination}.

In between sessions, we also  exchange links to interesting content on
a Slack channel.

\section{Implementation challenges}

Over the course of the last two years, we had to make some adjustments
to the way we carry out the  activity, an experience from which we are
also learning.

We reside  in different geographical  areas located in  different time
zones. The sessions take place online at a time which corresponds to morning or
late afternoon calls depending on our  location. The choice of a video
conferencing tool suitable for our requirements and needs  has not been
straightforward.   We want to use something that is easily  accessible, secure  and
respectful of privacy.

When we  started the  group, we did  not set any  agenda prior  to the
session.  The discussion  was  entirely unstructured.  We did  however
realize that such a format was  not sustainable in the long run. There
was naturally a need for some directions. We therefore decided to turn
the discussion room into a reading  group in which one person is to
present a  piece of  work to the  others, from  which a
discussion could emerge.

It  is essential  for us  to keep  the reading  group informal.  While
highly rewarding  from an  intellectual perspective, this  activity is
(currently)  a  side  project  and  we  believe  that  making  it  too
“academic” might impede  our enthusiasm. At the same  time, it rapidly
became evident  that setting a goal is  necessary to keep  the dynamic
going. Given that we work in  different domains and sectors, finding a
common target for this initiative can be challenging.

\section{Next steps}

The reading  group has  enabled us  to go  through many  different use
cases, reported  in the press,  in the scientific  literature, through
outreach  initiatives,  \textit{etc.}  We  aim to  pursue  this  initiative  in
different directions.

We wish to extend the sessions of the reading group to new participants. In
particular, we are interested in kicking off collaborations with practitioners
from other disciplines who have an interest in the topics of the café-philo.
In addition, we are interested in building on top an initiative from two
researchers in mathematics associated with the University of Cambridge in the
UK and the University of Aachen in Germany who have been implementing a
framework in the form of a comprehensive manifesto \cite{Chiodo23} to guide
the responsible and ethical development of data-driven projects. A co-author
of the paper is currently experimenting with using the framework in teaching
mobility data analytics with Python as part of a final year module at École
des Ponts ParisTech. The objective is to raise awareness among students of the
ethical challenges associated with developing a project that uses data to
support decision-making and make them think outside of a purely
performance-driven mindset. More generally, we wish to develop a framework
similar to the one proposed by Chiodo and Müller, adapted to the specifics
of computer science. To that end, we plan to perform a benchmark of existing
toolsets for teaching ethics to computer scientists, especially in French
Doctoral Schools (\textit{e.g.,} \cite{MoocEthics}). We would like to invite
anyone interested in this initiative to get in touch with us.


\bibliographystyle{alpha} \bibliography{paper}

%\vspace*{-\baselineskip}
\begin{IEEEbiographynophoto}{Marc Bruyère}
  began  his  career  in  1996  with  the  Internet  Service  Provider
  Club-Internet.   Over   the  years,  he  has   worked  with  various
  organizations,  including Cisco,  Vivendi  Universal, Credit  Suisse
  First  Boston, Airbus/Dimension  Data, Force10  Networks, and  Dell.
  His academic journey  started in 2012 with a Ph.D.  at the LAAS CNRS
  and  a two-year  PostDoc at  the University  of Tokyo.  His doctoral
  thesis  is  about open-source  OpenFlow  SDN  for Internet  Exchange
  Points (IXPs). Today, he is a senior researcher at IIJ Lab in Japan.

%  The rapid  evolution of  internet infrastructure  over the  past few
%  decades  has  undeniably transformed  the  landscape  of our  global
%  society.   With approximately  65\%  of the  world’s population  now
%  connected  online, the  pervasive nature  of this  digital realm  is
%  evident.  While this connectivity has brought many opportunities and
%  benefits, it has  also given rise to  significant challenges.  Among
%  these are the emerging societal  issues, such as the digital divide,
%  the society  of control that also  appeared as a unique  response to
%  the security  or control  dilemma, and the  undeniable environmental
%  implications and sustainability issues.
%
%  Given these challenges,  it becomes essential to  address a research
%  question:  How  can we  envision  Internet  and computer  networking
%  technologies  to  be  more   attuned  to  human-centric  values  and
%  prioritize small, localized scales of operation?
%
%  Computer  networking researchers  can  not ignore  or avoid  ethical
%  questions.   Marc has  been  examining and  exploring the  potential
%  avenue  in  low-tech  solutions  combined  with  free  software  and
%  hardware as  choices.  By advocating  a low-tech approach,  we could
%  potentially  reduce  the  environmental  footprint  of  our  digital
%  activities.   Using free  software and  hardware ensures  that these
%  technologies  remain accessible,  adaptable,  and resilient  against
%  monopolistic  tendencies that  can sometimes  stifle innovation  and
%  inclusivity.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{Florian Richoux}
  is  a senior  research at  the AI  Research Center  of the  National
  Institute of  Advanced Industrial  Science and Technology  in Tokyo,
  Japan.  His research focuses  on Combinatorial Optimization combined
  with Machine Learning, often applied on games.

% The  fruits of  AI research  can unlock  tremendous possibilities  and
% opportunities  for people,  but it  can also  be misused  for purposes
% going against  the common good.  Like the  Laws on Bioethics  voted in
% France in the  late 90’s, we need a legal  frame to delimit boundaries
% between what is  an ethical usage of  AI and what it  is not, defining
% what kind of AI usage we want for our societies.

% This is the purpose of the European AI Act that has been passed by the
% European Parliament on June 2023. This  act proposes a first set of AI
% regulations, but  may not cover all  possible usage of AI  yet: only 6
% months before  its adoption by  the European Parliament,  the European
% Council  adopted a  version of  the text  without taking  into account
% Generative AIs, such  as ChatGPT or DALL·E. As  we know, technological
% improvements in  AI, and in  particular in Machine Learning  (ML), are
% advancing rapidly,  and making difficult for  legislative apparatus to
% keep up.  It is our responsability,  as actors of AI  advancements, to
% keep ourselves  informed of the  new regulations around AI,  to detect
% and analyse  the gaps and  missing rules of  such acts, and  to inform
% both the people and policy makers about what remains to be done.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{Daphné Tuncer}
  has been working  in the academic research arena for  15 years.  Her
  research interests are in the  domain of network and computer system
  management.  While she  has mostly  been publishing  pieces of  work
  focusing on technical solutions until now, she has a strong interest
  in the socio-economics of the Internet and digitalization, and their
  impact on  people and the  environment. She reads about,  listens to
  and watches a  lot about ethics, history, and  geopolitics, that are
  all great sources of inspiration  to question research from the what
  for, why  and how perspective.   Daphné is currently with  École des
  Ponts ParisTech,  France. She  initiated the  reading group  in 2021
  with Marc Bruyère.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{Marc van der Wal}
  is an  R\&D engineer  at Afnic,  the registry  for domain  names for
  France (.fr) and overseas territories.

  Among his research and development projects, he is exploring ideas for
  fighting abuse (spam, phishing, etc.) involving domain names. While some
  registries already use machine learning-based techniques for that purpose,
  he wants to make sure that no ethical issues might arise from doing the same
  on .fr domain names. He is particularly interested in explainability of AI
  models, questions regarding privacy and combating digital exclusion.
\end{IEEEbiographynophoto}

\end{document}
